Symbolic linearization of constrained multibody systems

### 2013-06-01 PetersonGedeHubbard2012.tex
Fixed dimensions of P' in Discussion section.
Added dynamics overview and rewrote intro
Introduction still needs work to complete it. Dynamics overview probably needs
review, but is "complete".

\documentclass[smallcondensed]{svjour3}                     % onecolumn (standard format)
%\documentclass{article}
\RequirePackage{fix-cm}

@@

\title{A linearization procedure for constrained multibody systems}

\titlerunning{A linearization procedure for constrained multibody systems}        % if too long for running head

@@

\begin{abstract}
% Context
Many common mechanical systems have configuration or velocity constraints.
% Need  (what we have and what we want)
Analyses of such systems often require linearized forms of the motion
equations.  Although derivation of the nonlinear motion equations is well
understood, very little has been presented on the proper approach to
forming linearized versions which satisfy the constraints.
% Task
To address this issue, we developed a procedure for organizing the constraints
and motion equations, and subsequently linearizing these equations.
% Object of the document
In this paper we present this procedure to linearize the nonlinear motion
equations of constrained multibody systems, and illustrate it with a familiar
example, the rolling disc.
% Findings
We apply this procedure to a model of a bicycle that includes a closed
kinematic loop and four nonholonomic constraints and find that the resulting
symbolic linearized motion equations are identical to previously published
results.  The technique can be performed symbolically (enabling explicit
analytical forms of the linearized dynamics) or numerically.
% Conclusion
Motion equations for constrained multibody systems can now be linearized in a
systematic fashion, symbolically or numerically, thereby facilitating common
linear analyses (e.g., stability analysis and control system design).
% Perspectives
Efficiency of computer implementations of this procedure will be considered in
future studies.
\keywords{linearization, constrained multibody systems, symbolic dynamics, control}
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
Multibody systems, rigid or flexible, and particles often have constraints
which limit how they can move. In this paper, we will discuss 3 types of
constraints: configuration, velocity, and acceleration. Configuration
constraints limit the location or orientation of parts of the system, relative
to the external world or other parts of the system. Velocity constraints limit
the speeds at which the configuration can change, either from configuration
constraints (through time differentiation) or independent application. In this
paper, acceleration constraints refer to time differentiated velocity
constraints. Examples of multibody systems with constraints include closed
kinematic loops or objects rolling without slip.

Linear equations of motion are necessary for many different analyses.
Stability analysis and control system design are two possible scenarios.
Linearization of these constrained systems is not straightforward and presents
significant challenges. Computing the Jacobian matrix of the differential
equations - without considering the system's constraints - will give incorrect
results.

Currently, only a few approaches to linearizing constrained systems which have
been presented\cite{Kang2003}, \cite{Negrut2006}. A limitation of these
approaches is that they have been developed for equations of motion found using
Lagrange's method, which contain Lagrange multipliers. Kane's method
\cite{Kane1985} allows formulation of equations of motion for constrained
systems in which no multipliers are present. The dynamics of these systems can
be completely described by ordinary differential equations. This significantly
different formulation requires a different approach to linearization than what
existing options provide.

The goal of this paper is to establish concretely the first order relationships
(about an arbitrary point of linearization) between a selection of independent
coordinates and speeds and the time-derivatives of all coordinates and speeds
(both independent and dependent). The formalism presented here is generic
enough to cover most examples of time-varying constrained multibody systems
with arbitrary external inputs and arbitrary specified quantities.  The
procedure was created for systems whose equations of motion have been derived
with Kane's method. Although also applicable to dynamic system equations
formulated using other methods, it is restricted to systems which can be
completely described by a set of ODEs (i.e., DAEs needn't be solved).

We first present the linearization procedure, followed by some discussion of
the procedure, together with an example. The derivation of the procedure is
provided in Section \ref{sec:derivations}. Additionally, the electronic
supplementary material provides more examples (in more detail) and an
alternative ordering for the independent/dependent quantities.

%The complete configuration of the system is further assumed to be described by
%generalized coordinates $\bm{q}\in\bm{R}^n$ and its velocity completely
%described by generalized speeds $\bm{u}\in\bm{R}^o$, with $n \ne o$ in general.
%All coordinates and speeds, not only the independent ones, are represented by
%$\bm{q}$ and $\bm{u}$.  The system is assumed to be governed by the
%relationships described in Table \ref{table:assumptions}.  The system is
%described by $l + m + n + o$ equations which relate $2n + 2o + s + 1$
%quantities, which implies that $n - l + o - m + s + 1$ of these quantities are
%independent.  This paper discusses the first order relationship between these
%independent quantities ($n-l$ independent coordinates, $o-m$ independent
%speeds, $s$ control inputs, and time $t$) and the time derivatives of the $n$
%coordinates and $o$ speeds.
%
%The first three equations in Table \ref{table:assumptions} represent
%constraints derived purely from kinematic considerations at the configuration,
%velocity, and acceleration levels, respectively.  The velocity constraints may
%be nonholonomic or time differentiated holonomic.  The acceleration constraints
%can be time differentiated velocity constraints or kinematic acceleration
%constraints.  The fourth equation represents kinematic differential equations
%which relate $\bm{\dot{q}}$ to $\bm{u}$ and are linear in both these
%quantities.  The fifth equation represents the constraint-free nonholonomic
%dynamic differential equations~\cite{Kane1985}, and will necessarily be linear
%in $\bm{\dot{u}}$ and the inputs $\bm{r}$, but generally nonlinear in all other
%quantities.  This formalism is exceedingly generic; in practice, the structure
%of the functions in Table \ref{table:assumptions} can be leveraged to simplify
%the following calculations.  This will be discussed subsequently.
%
%Trajectories of the system exist on a $p \triangleq n - l + o - m$ dimensional
%manifold embedded in a $n + o$ dimensional space, though conserved quantities
%(e.g., total mechanical energy) may restrict trajectories to a lower
%dimensional manifold.  In general this cannot be assumed, however, especially
%in the presence of exogenous inputs $\bm{r}$ which may change the total
%mechanical energy of the system.  While more than $p$ of the independent state
%variables may be of interest (i.e., dependent coordinates or speeds may be of
%interest in themselves), only $p$ quantities are truly independent.  Which of
%these $p$ coordinates and speeds should be chosen as independent out of the
%possible $n + o$ may be obvious for small systems where there is intuition
%about system behavior, but in general the optimal choice depends on how the
%configuration $\bm{q}$ affects the gradients of $f_c$ and $f_v$.  On the other
%hand, in systems with cyclic (ignorable) coordinates, fewer than $p$ state
%variables may be of interest; one example is the Whipple bicycle model where
%$p=10$ but 5 of these state variables are cyclic coordinates (Cartesian
%coordinates of rear wheel contact point, heading, and both wheel angles) and
%may only be of secondary interest.

@@

\section{Discussion}
\label{sec:discussion}
Equation (\ref{eq:state_space_constrained}) doesn't fit into the standard linear
system framework because it is non-square.  To obtain a square system (useful
in stability analysis and control system design) we define
\begin{align}
  \label{eq:A_prime}
    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
  \label{eq:B_prime}
    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
\end{align}
where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
independent states by defining
\begin{align}
  \label{eq:P_prime}
    P^\prime &\triangleq \begin{bmatrix}
        P_{qi} & \bm{O}_{n \times (o - m)} \\
        \bm{O}_{o \times (n - l)} & P_{ui}
    \end{bmatrix} \\
  \label{eq:A}
    A &\triangleq P^{\prime T} A^\prime \\
  \label{eq:B}
    B &\triangleq P^{\prime T} B^\prime
\end{align}
where $P^\prime \in \bm{R}^{(o - m + n - l) \times (o - m + n - l)}$.  Defining
$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
linear systems analyses may applied.  It is worth noting that the rows
of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
in the output or measurement equations of a linear state space model, as in the
case when the controlling or measuring dependent state(s).

>>>>

\documentclass[smallcondensed,final]{svjour3}                     % onecolumn (standard format)
%\documentclass{article}
\RequirePackage{fix-cm}

@@

\title{Symbolic linearization of constrained multibody systems}

\titlerunning{Symbolic linearization of constrained multibody systems}        % if too long for running head

@@

\begin{abstract}
% Context
Many common mechanical systems have configuration or velocity constraints.
% Need  (what we have and what we want)
Analyses of such systems often require linearized forms of the motion
equations.
% Task
To address this issue, we developed a procedure for organizing the constraints
and motion equations, and subsequently linearizing these equations.
This procedure was specifically developed for equations of motion generated by
Kane's method, but is compatible with any method as long as the following are
true: only ordinary differential equations are required to describe the system,
and independent and dependent states remain in the final equations (i.e. no
DAEs, no substitution of variables).
% Object of the document
Following a brief review of Kane's method and the structure of the equations it
generates, we present the procedure to symbolically linearize the nonlinear
motion equations of constrained multibody systems, and illustrate it with the
familiar example of the rolling disc.

\keywords{linearization, constrained multibody systems, symbolic dynamics, control}
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
Designing a feedback controller or performing stability analysis are two common
analyses performed on models of multibody systems.
A linear form of the equations of motion is required to apply the most
frequently used techniques to such systems.
Normally, obtaining the equations of motion is as simple as taking the Jacobian
of the right hand side function of the equations of motion with respect to the
states.
However, when the multibody system in questions has constraints, this process
is more complicated.
If one has used Kane's method \cite{Kane1985} to generate the equations for a
constrained system, the resulting equations will have a particular form: a set
of ordinary differential equations containing both the independent and
dependent states.
Simple inspection of the equations does not reveal that they represent a
constrained system - their form is similar to the equations of an unconstrained
system.
It is critical to state that linearizing these equations by computing the
Jacobian will give incorrect results.
This paper provides a way to correctly generate the linear equations of motion
for constrained systems which have been found using Kane's method (or any other
method which results in equations with the same form).

Multibody systems, rigid or flexible, and particles often have constraints
which limit how they can move or be configured. In this paper, we will discuss
three types of constraints: configuration, velocity, and acceleration.
Configuration constraints limit the location or orientation of parts of the
system, relative to the external world or other parts of the system. Velocity
constraints limit the speeds at which the configuration can change, either from
configuration constraints (through time differentiation) or independent
application. In this paper, acceleration constraints refer to time
differentiated velocity constraints. Motion constraints most often appear in
systems where there is rolling without slip or there are closed kinematic
loops.

% TODO: Improve literature review.
Bottema appears to be the first author to show that special considerations are
needed for linearizing non-holonomic systems \cite{Bottema1949}.
While other authors have examined linearization of nonholonomic systems, we
have found isssues with applying any of them to equations generated by Kane's
method.
Kang et al. \cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both
explored linearization, but these methods have been developed for equations of
motion found using Lagrange's method, which contain Lagrange multipliers (which
are not present in equations of motion generated from Kane's method).
Minaker and Rieveley \cite{Minaker2010} and Schwab and Meijaard
\cite{Schwab2003} both have developed techniques for generating equations of
motion (and linear forms), however both techniques are more numerical in
nature - they do not provide symbolic equations of motion or symbolic
linearized equations.
These processes, along with others such as Neuman's %TODO: fix reference
, formulate the linear equations as a step in the equations of motion
generation process.
They are not capable of linearizing existing equations of motion.

% TODO:reiterate limiations of model
The goal of this paper is to establish concretely the first order relationships
(about an arbitrary point of linearization) between a selection of independent
coordinates and speeds and the time-derivatives of all coordinates and speeds
(both independent and dependent). The formalism presented here is generic
enough to cover most examples of time-varying constrained multibody systems
with arbitrary external inputs and arbitrary specified quantities.  The
procedure was created for systems whose equations of motion have been derived
with Kane's method. Although also applicable to dynamic system equations
formulated using other methods, it is restricted to systems which can be
completely described by a set of ODEs (i.e., DAEs needn't be solved).

% TODO:rewrite organization of paper
The use of Kane's method in generating equations of motion is reviewed first,
in order to properly orient the reader to the format and some of the qualities
of the generated equations.
Kane's method is then briefly used to generate the equations of motion for a
simple system, a rolling disk with a trailing arm.
The need for a different linearization technique is also demonstrated with this
model.
In response to this need, we present the newly developed linearization
procedure, followed by its derivation. % maybe flip the order?
This new linearization procedure is applied to the example system, to provide
an example of its application.
Finally, we discuss other details and nuances of the procedure and its use.

\section{Kane's Method, Briefly}
\label{sec:kane_method}
To familiarize the reader with Kane's method, three concepts relating to its
use will be examined.
The first is the use of generalized speeds in addition to generalized
coordinates.
The second concept is in how these generalized speeds can be used to project
permissible motions of the system, and how this removes the need to consider
non-contributing (internal constraint) forces.
Finally, the mathematical steps required to use Kane's method are shown and
described.

In using Lagrange's method, generalized coordinates are used to describe the
configuration of a system while the time derivatives of the gen. coordinates
describe the velocities of a system.
Kane's method allows for the velocities of a system to be written in terms of
generalized speeds, which are not required to be the time derivatives of
coordinates (although such a definition is permitted).

The benefit of using both generalized speeds and coordinates can be seen in
the following example.
If the orientation of a rigid body were to be described using using Euler
parameters (quaternions), its angular velocity is relatively complicated when
using the time derivatives of the coordinates.
Using generalized speeds allows for the angular velocity to be written in a
much simpler form:
\begin{align}
\label{eq:ex1}
{^N}\bar{\omega}^B = \it{\omega}_1 \hat{b}_1 + \it{\omega}_2 \hat{b}_2 +
\it{\omega}_3 \hat{b}_3
\end{align}
or, the angular velocity of rigid body $B$ in the reference frame $N$ is the
defined as the sum of three generalized-speed/basis-vector products.
The derivatives of these generalized speeds will then appear in the equations
of motion, rather than the twice time differentiated generalized coordinates.

A downside of allowing such definitions of system velocities is that the
kinematic differential equations, which relate the rate of change of the gen.
coordinates to the gen. speeds, become more complicated.
This penalty is usually offset by the vastly simplified equations of motion
that are generated.

Using and understanding Kane's method requires another concept: partial
velocities.
Partial velocities are defined as the partial derivatives of the velocity
vectors with respect to the generalized speeds.
They can be defined for both translational and rotational velocities.
Using the angular velocity of body $B$ in reference frame $N$ as shown in
(\ref{eq:ex1}), the partial velocities can be written as:
\[
{^N}\bar{\omega}^B_{\omega_1} = \hat{b}_1 \quad \quad
{^N}\bar{\omega}^B_{\omega_2} = \hat{b}_2 \quad \quad
{^N}\bar{\omega}^B_{\omega_3} = \hat{b}_3
\]
The partial velocities represent the direction of motion associated with each
generalized speed.

By taking the dot product of each partial velocity and both sides of Newton's
second law, $F=ma$ (or the rotational equivalent), only terms that are
parallel to the partial velocities are left in the equation.
One equation is generated for each generalized speed.

An important byproduct of this is that internal constraint forces
(non-contributing forces) generated by things such as pin or sliding joints no
longer need to be considered.
Assuming the generalized speeds have been defined in an intelligent manner, the
forces involved in the internal constraints cannot affect the rate of change of
any generalized speed, and thus do not need to be considered.

In summary, Kane's method allows for velocities to be defined in terms of
generalized speeds, allowing for simpler velocity expressions.
The dot product of each partial velocity (the direction of motion associated
with each generalized speed) with Newton's second law (or Euler's equations)
removes all terms which are not related to each generalized speed.
This generally results in simple equations of motion, which are relatively easy
to form.

The mathematical formalism for applying Kane's method to multibody systems
follows.
Consider a system composed of rigid bodies $B_1,...,B_g$, particles
$P_1,...,P_h$, with points of interest $Q_1,...,Q_k$, and reference frames of
interest $E_1,...,E_c$, all defined relative to the inertial reference frame
$N$.
This system has generalized coordinates $q_1,...,q_n$, generalized speeds
$u_1,...,u_o$.
Additionally, there are $l$ configuration constraints and $m$ velocity
constraints.

In order to properly apply Kane's method, the velocities of each component in
the system need to be written in the following form.
Translational velocity of particles:
\begin{align}
\label{eq:particle_translational}
{^N}\bar{v}^{P_i} = \sum^o_{j=1} {^N}\bar{v}^{P_i}_{u_j} u_j + {^N}\bar{v}^{P_i}_t
\quad \quad i=1,...,g
\end{align}
Translational velocity of rigid body mass centers:
\begin{align}
\label{eq:rb_translational}
{^N}\bar{v}^{B_i^*} = \sum_{j=1}^o {^N}\bar{v}^{B_i^*}_{u_j} u_j +
{^N}\bar{v}^{B_i^*}_t \quad \quad i=1,...,h
\end{align}
Translational velocity of points of interest:
\begin{align}
\label{eq:points_translational}
{^N}\bar{v}^{Q_i} = \sum_{j=1}^o {^N}\bar{v}^{Q_i}_{u_j} u_j + {^N}\bar{v}^{Q_i}_t
\quad \quad i=1,...,k
\end{align}
Rotational velocity of rigid bodies:
\begin{align}
\label{eq:rb_rotational}
{^N}\bar{\omega}^{B_i} = \sum_{j=1}^o {^N}\bar{\omega}^{B_i}_{u_j} u_j +
{^N}\bar{\omega}^{B_i}_t \quad \quad i=1,...,h
\end{align}
Rotational velocity of reference frames of interest:
\begin{align}
\label{eq:rb_rotational}
{^N}\bar{\omega}^{E_i} = \sum_{j=1}^o {^N}\bar{\omega}^{E_i}_{u_j} u_j +
{^N}\bar{\omega}^{E_i}_t \quad \quad i=1,...,c
\end{align}

In these equations, each term $\bar{v}_t$ or $\bar{\omega}_t$ represents the
velocity or angular velocity component which is a prescribed function of time,
and has no dependence on generalized coordinates or speeds (e.g., a crank
driven at a fixed rate).

For this system, the configuration constraints are defined as:
\begin{align}
\label{eq:configuration_constraints}
\mathbf{f}_c(\mathbf{q}, t) = \mathbf{0}
\end{align}
where the terms $\mathbf{f}_c$ and $\mathbf{0}$ have the shape $l \times 1$.
The velocity constraints are defined as:
\begin{align}
\label{eq:velocity_constraints}
\mathbf{f}_v(\mathbf{q}, \mathbf{u}, t) = \mathbf{0}
\end{align}
where the terms $\mathbf{f}_v$ and $\mathbf{0}$ have the shape $m \times 1$.
It is important that the velocity constraints include the time-differentiated
configuration constraints or equivalent constraints which produce the same
behavior.
Furthermore, velocity constraints can not be written with $\dot{\mathbf{q}}$
terms (the time derivatives of the coordinates).
Acceleration constraints need to be defined as well.
Typically, they will simply by time-differentiated velocity constraints and
have the form:
\begin{align}
\label{eq:acceleration_constraints}
\mathbf{f}_a(\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, \dot{\mathbf{u}}, t) =
\mathbf{0}
\end{align}

At this point, it is convienient to define a matrix $\mathbf{B}$, which is the
gradient of the velocity constraints $\mathbf{f}_v$ with respect to all of the
generalized speeds $\mathbf{u}$.
\begin{align}
\label{eq:constraint_B}
\mathbf{B} \triangleq \nabla_{\mathbf{u}} \mathbf{f}_v (\mathbf{q}, \mathbf{u},
t)
\end{align}
and from (\ref{eq:velocity_constraints}) and (\ref{eq:constraint_B}) the
following can be written:
\begin{align}
\label{eq:constraint_Bu0}
\mathbf{0} = \mathbf{B}\mathbf{u}
\end{align}

With these definitions established, Kane's method can be applied.
The output of Kane's method is the equation $\mathbf{F} + \mathbf{F}^* =
\mathbf{0}$ - the sum of the generalized active forces and generalized inertial
forces is 0.
If Newton's second law were written as $f - ma = 0$, the $\mathbf{F}$ is the
multibody generalization of $f$ and $\mathbf{F}^*$ is the multibody
generalization of $-ma$.
These terms are created by taking the dot product of the partial
velocities/angular-velocities, as defined in (\ref{eq:particle_translational})
- (\ref{eq:rb_rotational}), with the resultant forces/torques, $\bar{R}$/
$\bar{T}$, and the inertial forces/torques, $\bar{R}^*$/$\bar{T}^*$.

The resultant force at a point is defined as the sum of all force vectors
acting directly on that point.
The resultant torque on a reference frame is defined similarly - it is the sum
of all torque vectors acting on that reference frame.
Note that distance forces do not need to be transformed into a resultant
force/torque for any components of the system; they only need to be considered
at the point of application (e.g., if a force $\bar{f}$ is the only force
applied to body $B$ and is applied at point $Q$, then $\bar{R}_{B^*}=0$ and
$\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$).
These resultant forces and torques are used to write the generalized active
forces:

\begin{equation}
\label{eq:definition_F}
\mathbf{F} =
\begin{bmatrix}
\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_1 +
\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_1 +
\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_1 +
\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_1 +
\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_1 \\
\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_2 +
\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_2 +
\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_2 +
\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_2 +
\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_2 \\
\displaystyle \vdots \\
\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_o +
\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_o +
\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_o +
\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_o +
\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_o
\end{bmatrix}
\end{equation}

The inertia forces and torques, $\bar{R}^*$ and $\bar{T}^*$, will now be
defined. Note: the reference frame these are defined is not written - at this
point in the derivation of the equations of motion, they are assumed to be in
the inertial frame, represented here as $N$.

Particles:
\begin{align}
\label{eq:particle_gen_inertia}
\bar{R}^*_P \triangleq -m_p {^N}\bar{a}^P
\end{align}
Rigid Bodies:
\begin{align}
\label{eq:rb_translational_gen_inertia}
\bar{R}^*_{B^*} \triangleq -m_B {^N}\bar{a}^{B^*}
\end{align}
\begin{align}
\label{eq:rb_rotational_gen_inertia}
\bar{T}^*_B \triangleq -{^N}\bar{\alpha}^B \times \bar{\bar{I}}^{B/B^*} -
{^N}\bar{\omega}^B \times \bar{\bar{I}}^{B/B^*} \cdot {^N}\omega^B
\end{align}
Where $\bar{\bar{I}}^{B/B^*}$ is the central inertia dyadic of the rigid body
$B$ about its mass center $B^*$ \ref{Kane1985}.

If the basis vectors of the reference frame associated with the rigid body $B$
are aligned with the principal axes of $B$, then the following definitions can
be used:
\begin{align}
\alpha_i \triangleq {^N}\alpha^B \cdot \hat{b}_i \quad \quad
\omega_i \triangleq {^N}\omega^B \cdot \hat{b}_i \quad \quad
I_i \triangleq \hat{b}_i \cdot \bar{\bar{I}}^{B/B^*} \cdot \hat{b}_i
\quad \quad for i=1,2,3
\end{align}
In this case, $I_1$, $I_2$, and $I_3$ will be the principal inertias of rigid
body $B$.
This allows the inertia torque to be written as:
\begin{align}
\label{eq:simple_rb_rotational_gen_inertia}
\bar{T}_B^* = -[\alpha_1 I_1 - \omega_2 \omega_3 (I_2 - I_3)] \hat{b}_1
              -[\alpha_2 I_2 - \omega_3 \omega_1 (I_3 - I_1)] \hat{b}_2
              -[\alpha_3 I_3 - \omega_1 \omega_2 (I_1 - I_2)] \hat{b}_3
\end{align}

Now $\mathbf{F}^*$ can be constructed in a manner similar to $\mathbf{F}$.

\begin{equation}
\label{eq:definition_Fstar}
\mathbf{F} =
\begin{bmatrix}
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_1 +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_1 +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_1 \\
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_2 +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_2 +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_2 \\
\displaystyle \vdots \\
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_o +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_o +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_o
\end{bmatrix}
\end{equation}

Now the equations $\mathbf{F} + \mathbf{F}^* = \mathbf{0}$ are defined.
Bu what has been written is valid only if there are no motion constraints.
The equations need to be transformed in order to not violate the behaviors
defined by the constraints.

Kane's Method allows for the equations of motion to be written using both
independent and dependent speeds, $\mathbf{u}_i$ and $\mathbf{u}_d$, without
the need for Lagrange multipliers.
Even though both independent and dependent speeds are in the final equations, a
choice still needs to be made as to which speeds will be the constrained
speeds.
The methods for doing this are outside the scope of this paper, but it is
important that whatever selection is made avoids singularities.

To simplify notation, it is assumed that the generalized speeds are already
ordered in a favorable manner:
\begin{align}
\label{eq:speeds_independent_dependent}
\mathbf{u} =
\begin{bmatrix}
\mathbf{u}_i \\
\mathbf{u}_d
\end{bmatrix}
\end{align}
The matrix $\mathbf{B}$ from (\ref{eq:constraint_B}) needs to be rewritten as
well:
\begin{align}
\label{eq:constraint_B_independent_dependent}
\mathbf{0} &= \mathbf{B} \mathbf{u}\\
\mathbf{0} &= \begin{bmatrix}
\mathbf{B}_i & \mathbf{B}_d
\end{bmatrix} \begin{bmatrix}
\mathbf{u}_i \\ \mathbf{u}_d
\end{bmatrix}
\end{align}
Where $\mathbf{B}_i$ has the shape $m \times o - m$ and $\mathbf{B}_d$ has the
shape $m \times m$. This equation can be rewritten as:
\begin{align}
\label{eq:constraint_A}
\mathbf{B}_i \mathbf{u}_i + \mathbf{B}_d \mathbf{u}_d = \mathbf{0} \\
\mathbf{B}_d \mathbf{u}_d = - \mathbf{B}_i \mathbf{u}_i \\
\mathbf{u}_d = -\mathbf{B}_d^{-1} \mathbf{B}_i \mathbf{u}_i \\
\mathbf{A} \triangle - \mathbf{B}_d^{-1} \mathbf{B}_i \\
\mathbf{u}_d = \mathbf{A} \mathbf{u}_i
\end{align}
Where $\mathbf{A}$ has the shape $m \times m$.

Finally, to constrain the generalized active and inertia forces, they need to
be written as:
\begin{align}
\label{eq:F_Fstar_rewrite}
\mathbf{F} = \begin{bmatrix} \mathbf{F}_i \\ \mathbf{F}_d \end{bmatrix} \quad
\quad \mathbf{F}^* = \begin{bmatrix} \mathbf{F}^*_i \\ \mathbf{F}^*_d
\end{bmatrix}
\end{align}
allowing the nonholonomic generalized forces to be written:
\begin{align}
\label{eq:F_Fstar_nonholonomic}
\tilde{\mathbf{F}} = \mathbf{F}_i + \mathbf{A}^T \mathbf{F}_d \quad \quad
\tilde{\mathbf{F}}^* = \mathbf{F}_i^* + \mathbf{A}^T \mathbf{F}_d^*
\end{align}

After constraining the equations of motion generated, we have reduced the
number of equations from $o$ to $o-m$; the equations still contain $o$ unknown
terms that need to be solved for ($\dot{\mathbf{u}}_1,...,\dot{\mathbf{u}}_o$).
In order to solve these second order differential equations for the
$\dot{\mathbf{u}}$ terms, additional equations are needed.
The acceleration constrain equations fill this role, as there are $m$ of these
equations, giving:
\begin{align}
\label{eq:F_Fstar_f_a}
\begin{bmatrix}
\tilde{\mathbf{F}} + \tilde{\mathbf{F}}^* \\
\mathbf{f}_a (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, \dot{\mathbf{u}}, t)
\end{bmatrix} = \mathbf{0}
\end{align}
which gives a total of $o$ equations, enough to be able to solve for the
$\dot{\mathbf{u}}$'s.

As a final step, the arrangement of the equations as used in this paper is
shown in Table \ref{table:assumptions}.
These equations are all drawing from the preceding steps, but have been
re-organized for convenience (as will be shown later).

The only new terms are $\mathbf{f}_0$, $\mathbf{f}_1$, $\mathbf{f}_2$, and
$\mathbf{f}_3$.
The kinematic differential equations are split up into the quantities
$\mathbf{f}_0$ and $\mathbf{f}_1$, depending on whether each term has a
$\mathbf{u}$ or $\dot{\mathbf{q}}$ present.
The same is true for $\mathbf{f}_2$ and $\mathbf{f}_3$; if a term contains a
$\dot{\mathbf{u}}$, it is placed into $\mathbf{f}_2$ and if not it is placed
into $\mathbf{f}_3$.

\section{Dynamics Example}
TODO: Luke writes here?

@@

\section{Discussion}
\label{sec:discussion}
Equation (\ref{eq:state_space_constrained}) doesn't fit into the standard linear
system framework because it is non-square.  To obtain a square system (useful
in stability analysis and control system design) we define
\begin{align}
  \label{eq:A_prime}
    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
  \label{eq:B_prime}
    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
\end{align}
where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
independent states by defining
\begin{align}
  \label{eq:P_prime}
    P^\prime &\triangleq \begin{bmatrix}
        P_{qi} & \bm{O}_{n \times (o - m)} \\
        \bm{O}_{o \times (n - l)} & P_{ui}
    \end{bmatrix} \\
  \label{eq:A}
    A &\triangleq P^{\prime T} A^\prime \\
  \label{eq:B}
    B &\triangleq P^{\prime T} B^\prime
\end{align}
where $P^\prime \in \bm{R}^{(o + n) \times (o - m + n - l)}$.  Defining
$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
linear systems analyses may applied.  It is worth noting that the rows
of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
in the output or measurement equations of a linear state space model, as in the
case when the controlling or measuring dependent state(s).
